---
title: "Data Cleaning and Standards"
output: html_notebook
author:
- Lina Aragón-Baquero
date: "`r format(Sys.time(), '%d %B %Y')`"
---

# **Selected Task**

For this first assignment, I will do the cleaning task #1: Convert all dates and geographic coordinates to ISO standards. [suggested packages: lubridate and rgdal; see example code in tutorial 2]. See ISO 8601 (Links to an external site.) for dates (yyyy-mm-dd). See ISO 6709 (Links to an external site.) for coordinates (UTM).

# **Description of my data cleaning task**

I choose to work with coordinates and dates because I used to find these two formats annoying and hard to work with. I used some of the code provided in tutorials 2 and 3 to upload the files in the BWG database. After that, I divided the task in two parts. First, I transform the coordinates to UTM system by transforming them to Spatial Points and the to the UTM coordinate system. Because I couldn’t create a continuous process to do this transformation and directly assign the new values to the “lng” and “lat” columns, I used the function mutate to replace the values in “lng” and “lat” with the values saved in the “xt_utm” matrix. The second part was to transform the two columns with dates (“bwg_release” and “public_release”) from "%d/%m/%Y" to "%Y/%m/%d". The important step here was to include in the code the type of format the column has because without providing it the process failed. Finally, I export the clean file as a .csv and saved in a folder inside my project called CleanData.

# **Script**

## 1. Load packages and database

```{r}
getwd()

#required packages
pkgs <- c("taxize", "assertr", "stringdist", 
          "tidyverse", "palmerpenguins", "GGally",
          "lubridate","rgdal","myTAI","dplyr","ggplot2","sf")

#Load all the required packages
lapply(pkgs, library, character.only = TRUE)

#Check files in the BWG database
BWGfiles <- list.files(path = "BWG_database/", pattern = "*.csv", full.names = TRUE)
BWGfiles

#Load all the files in the BWG database
list2env(
  lapply(
    setNames(BWGfiles, 
             make.names(
               gsub(".*1_", "", 
                    tools::file_path_sans_ext(BWGfiles)))), 
    read_csv), 
  envir = .GlobalEnv)

```

## 2. Convert coordinates to ISO in the the file "datasets" using code from tutorials

```{r}
#A. Summary datasets
head(datasets)
str(datasets)

#B. Extract latitude and longitude
xy<-datasets[c("lng","lat")]
xy

#B. Convert them to UTM
## 1. Convert coordinates to "Spatial Points"

coordinates(xy)<-c("lng","lat")
proj4string(xy)<-CRS("+proj=longlat +datum=WGS84")
xy

##2. Transform xy "Spatial Points" to UTM coordinate system
xy_utm<-spTransform(xy, CRS("+proj=utm +zone=16 +datum=WGS84"))
xy_utm


#D. Change the values of Latitude and Longitude in the datasets file and check
datasets<- datasets %>% mutate(lng=xy_utm@coords[,1],lat=xy_utm@coords[,2])
head(datasets)
```


## 3. Transform bwg_release and public_release to the correct date format

```{r}
#1. Check the formtat of the columns of interest

class(datasets$bwg_release) #character
class(datasets$public_release) #character

#2. Transform the columns to type "Date"
datasets$bwg_release <- lubridate::as_date(datasets$bwg_release, format = "%d/%m/%Y")
class(datasets$bwg_release) 

datasets$public_release <- lubridate::as_date(datasets$public_release, format = "%d/%m/%Y")
class(datasets$bwg_release) 

```

## 4. Check and exports the new datasets files as .csv. Choose the folder CleanData as the designated location to keep track of the different versions. 

```{r}
head(datasets)
write_csv(datasets, "CleanData/datasets.csv")
```




